Hospital Information Technology and Physician Response

### 2021-10-15 Presentations/ThirdYearPaperpres.tex
Updates from Overleaf

\usetheme[progressbar=frametitle]{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage[style=authoryear]{biblatex}
\bibliography{demo.bib}
\usepackage{pgfplots}

@@

\begin{frame}{The Potential of HIT}
\centering
    \includegraphics[scale=.6]{graphics/News Clip4.PNG}
\end{frame}

@@

\begin{frame}{Physician's Incentives}
\begin{itemize}
    \item Physician decides to work, possibly at a handful of hospitals
    \item Some proportion of these implement a new technology (an EHR)
    \item Two ways to think about the effect:
    \begin{enumerate}
        \item The cost of switching hospitals is higher (attachment to EHR hospital)
        \item The cost of switching hospitals is lower (increased burden at EHR hospital)
    \end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}{Data}
\begin{equation*}
    \textcolor{blue}{p_{iht}}=f(EHR_{ht},EHR_{-h,t},X_i,H_{jt},...)
\end{equation*}

\vspace{3mm}

Measures of physician labor market behavior
\begin{itemize}
    \item Number/percent of shared patients physician $i$ has with hospital $h$: \underline{CMS Shared Patient Data} (2009-2015)
    \begin{itemize}
        \item Main dataset: physician hospital level panel
        \item Physicians and hospitals in this data work closely together, likely the physician works \textit{in} the hospital
    \end{itemize}
    \item Total Part B claims for physician $i$: \underline{Medicare Provider Utilization and Payment Data} (starts in 2012)
\end{itemize}
\end{frame}

\begin{frame}{Data}
\begin{equation*}
    p_{iht}=f(\textcolor{blue}{EHR_{ht},EHR_{-h,t}},X_i,H_{jt},...)
\end{equation*}

@@

\begin{itemize}
    \item Main variable (binary for whether a hospital uses an EHR) comes from AHA survey
    \item AHA IT Supplement also has information on EHRs that can help with documentation or decision making
    \item (\textit{Considering Later}) Meaningful use data gives information on which hospitals received a subsidy for achieving "meaningful use"
\end{itemize}
\end{frame}

\begin{frame}{Data}
\begin{equation*}
    p_{iht}=f(EHR_{ht},EHR_{-h,t},\textcolor{blue}{X_i},H_{jt},...)
\end{equation*}

Physician level characteristics
\begin{itemize}
    \item Medical school graduation date, gender: \underline{Physician Compare}
\end{itemize}
\end{frame}

\begin{frame}{Data}
\begin{equation*}
    p_{iht}=f(EHR_{ht},EHR_{-h,t},X_i,\textcolor{blue}{H_{jt}},...)
\end{equation*}

Hospital Level Characteristics
\begin{itemize}
    \item Number of beds: AHA Survey
    \item Number of days out of the year operating: AHA Survey
\end{itemize}
\end{frame}

\begin{frame}{Summary Statistics}
\centering

\includegraphics[scale=.8]{Objects/sumstats_pair_table.pdf}

\vspace{3mm}

\includegraphics[scale=.8]{Objects/sumstats_physician_table.pdf}

\begin{frame}{Summary Statistics}
\centering
\includegraphics[scale=.8]{Objects/sumstats_hospital_table.pdf}

\begin{frame}{Variation in EHR Adoption}
\centering
\includegraphics[scale=.53]{Objects/TYP_plot_hospEHR_year.pdf}
\end{frame}

\begin{frame}{Looking at Shares when EHRs are Adopted}
    \centering
    \includegraphics[scale=.5]{Objects/relyear_treatment.pdf}
\end{frame}

\section{Plans}

\begin{frame}{Initial Plan: Difference in Difference}
\begin{itemize}
    \item Interested in two treatment effects
    \begin{enumerate}
        \item The effect of EHR at hospital $h$ on share at hospital $h$
        \item The effect of EHR at hospital $h$ on share at hospital $-h$
    \end{enumerate}

    \vspace{2mm}

    \item My idea is to estimate these two separately, but there are issues because of spillover effects (hospital $-h$ implements an EHR)

    \vspace{2mm}

    \item Control group will be physicians who have not yet been exposed to an EHR at \textbf{any} hospital at time $t$
\end{itemize}
\end{frame}

\begin{frame}{Other Ideas}
\begin{itemize}
    \item Aggregate to physician level dataset if there is insufficient data for the control group
    \item Have a continuous measure of EHR use that captures share of exposure to EHRs
    \item IV: bandwidth, state privacy laws
\end{itemize}
\end{frame}

\begin{frame}{Other Points}
\begin{itemize}
    \item Want to dig deeper into how a hospital qualified for meaningful use subsidies
    \item Physician location
    \item Bring in other measures of work from shared patient data
\end{itemize}

>>>>

\usetheme[progressbar=frametitle]{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage[style=authoryear]{biblatex}
\bibliography{demo.bib}
\usepackage{pgfplots}
\usepackage{bbm}

@@

\begin{frame}{What is an EHR?}

\begin{itemize}
    \item Computer-program technology that hospitals purchase
    \item Performance of EHR depends on level purchased:
    \begin{itemize}
        \item All include electronic patient records
        \item More complex systems allow for decision making assistance
        \item More recently, patient interfaces are being developed
    \end{itemize}
\end{itemize}

@@

\begin{frame}{Physician's Incentives}
\begin{itemize}
    \item Physician has a choice set of where to work, including multiple hospitals and office-based locations
    \item All physicians I consider work closely with at least one hospital at some point in time
    \item Some proportion of hospitals implement a new technology (an EHR)
    \item How to think about the effect on labor:
    \begin{enumerate}
        \item EHR makes seeing patients easier (or more difficult), thus the number of patients seen increases (or decreases)
        \item EHR imposes a burden with a high cost, so the physician switches to working at an office-based location where they have more control
        \item A senior physician could be tipped to retire early instead of taking on the cost of learning a new technology
    \end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}{Measures of Physician Labor Market Decisions}

     Total Patients: \underline{CMS Shared Patient Data} (2009-2015)
    \begin{itemize}
        \item Sum of shared patients with all entities that have an NPI
    \end{itemize}

     Indicator for whether they are working
    \begin{itemize}
        \item =1 if Total Patients $>0$
    \end{itemize}

     Indicator for working in office-based location
    \begin{itemize}
        \item Comes from SK$\&$A database of office-based physicians
    \end{itemize}

\begin{frame}{Measures of EHR Use}

Physician-level Indicator that captures exposure to \textit{any} EHR
\begin{itemize}
    \item With information on shared patients between physicians and hospitals, I limit to physicians working in hospitals
    \item Fraction of hospitals worked with that use an EHR comes from the AHA survey
    \item Similarly, I can capture hospitals that use EHRs for decision making vs. documentation using the AHA Survey IT Supplement 
\end{itemize}

\begin{frame}{Other Characteristics}

Medical school graduation date, gender: \underline{Physician Compare}

@@

Average size of hospital worked with (measured in beds), average hospital days operating: \underline{AHA Survey}
\end{frame}

\section{Analysis}

\begin{frame}{Event Studies}
\begin{equation*}
    Y_{it}= \sum_{t=-6}^6 \mathbbm{1}_{\{\text{relative treatment year}_{i}=t\}} + \alpha_t + \delta_{i} + \mathbb{X}_{it}
\end{equation*}
